---
title: "Prediction of Manner via Machine Learning"
author: "Jerome Zhao"
date: "Monday, July 20, 2015"
output: html_document
---

## Executive summary

This project predicts the manner in which people do exercise. This is the "classe" variable in the training set. The prediction is based on the training data set and validated against the testing data set.

## Data preparation and loading

Load data from file system and show the class distribution. Basically all classes are equally possible.

```{r message=FALSE}
raw_data.train <- read.csv('pml-training.csv', na.strings = c('NA', ''), header = TRUE)
raw_data.test <- read.csv('pml-testing.csv', na.strings = c('NA', ''), header = TRUE)
library(caret)
library(randomForest)
set.seed(123)

barplot(table(raw_data.train$classe), main = 'Class distribution')
```


#### Setting predictors

Choose predictors which are relevant for the class prediction:

```{r}
predictors <- c('roll_belt', 'pitch_belt', 'yaw_belt', 'total_accel_belt',
                          'gyros_belt_x', 'gyros_belt_y', 'gyros_belt_z',
                          'accel_belt_x', 'accel_belt_y', 'accel_belt_z',
                          'magnet_belt_x', 'magnet_belt_y', 'magnet_belt_z',
                          'roll_arm', 'pitch_arm', 'yaw_arm', 
                          'gyros_arm_x', 'gyros_arm_y', 'gyros_arm_z', 
                          'accel_arm_x', 'accel_arm_y', 'accel_arm_z',
                          'magnet_arm_x', 'magnet_arm_y', 'magnet_arm_z',
                          'roll_dumbbell', 'pitch_dumbbell', 'yaw_dumbbell',
                          'gyros_dumbbell_x', 'gyros_dumbbell_y', 'gyros_dumbbell_z',
                          'accel_dumbbell_x', 'accel_dumbbell_y', 'accel_dumbbell_z',
                          'magnet_dumbbell_x', 'magnet_dumbbell_y', 'magnet_dumbbell_z',
                          'roll_forearm', 'pitch_forearm', 'yaw_forearm',
                          'gyros_forearm_x', 'gyros_forearm_y', 'gyros_forearm_z',
                          'accel_forearm_x', 'accel_forearm_y', 'accel_forearm_z',
                          'magnet_forearm_x', 'magnet_forearm_y', 'magnet_forearm_z',
                          'classe')

raw_data.train <- raw_data.train[,predictors]
raw_data.test <- raw_data.test[,predictors[-50]]
```



#### Data splitting

Data is split into two data sets: for training (60% of the data) and for validation (40%).

```{r}
inTrain <- createDataPartition(raw_data.train$classe, p = 0.6, list = FALSE)
training <- raw_data.train[inTrain,]
validation <- raw_data.train[-inTrain,]
```


## Prediction model training

#### Radnom Forest Model

"Random Forest" modelis quite efficient with a large number of predictors.

```{r}
modFit <- randomForest(classe ~ ., data = training)
modFit
```

#### Choose the 10 most important predictors

```{r}
varImpPlot(modFit, n.var = 10, main = '10 Most Important Predictors')
```

## Validation

```{r}
predictions <- predict(modFit, validation)
confusionMatrix(predictions, validation$classe)
```

#### Out of sample error

```{r}
confusionMatrix(predictions, validation$classe)$overall['Accuracy']
```

Very good accuracy

## Prediction on the test set

```{r}
predict(modFit, raw_data.test)
```

